% T I T L E   P A G E
% -------------------
% Last updated May 24, 2011, by Stephen Carr, IST-Client Services
% The title page is counted as page `i' but we need to suppress the
% page number.  We also don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
        \begin{center}
        \vspace*{1.0cm}

        \Huge
        {\bf Approximately Optimum Search Trees in External Memory Models }

        \vspace*{1.0cm}

        \normalsize
        by \\

        \vspace*{1.0cm}

        \Large
        Oliver Grant \\

        \vspace*{3.0cm}

        \normalsize
        A thesis \\
        presented to the University of Waterloo \\ 
        in fulfillment of the \\
        thesis requirement for the degree of \\
        Master of Mathematics \\
        in \\
        Computer Science \\

        \vspace*{2.0cm}

        Waterloo, Ontario, Canada, 2016 \\

        \vspace*{1.0cm}

        \copyright\ Oliver Grant 2016 \

        \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage % Ends the current page and causes all figures and tables that have so far appeared in the input to be printed.
% In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.
 


% D E C L A R A T I O N   P A G E
% -------------------------------
  % The following is the sample Delaration Page as provided by the GSO
  % December 13th, 2006.  It is designed for an electronic thesis.
  \noindent
I hereby declare that I am the sole author of this thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

  \bigskip
  
  \noindent
I understand that my thesis may be made electronically available to the public.

\cleardoublepage
%\newpage

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

We examine optimal and near optimal solutions to the classic binary search tree problem of Knuth. We are given a set of $n$ keys (originally known as words), $B_1, B_2, ..., B_n$ and $2n+1$ frequencies. ${p_1, p_2, ..., p_n}$ represent the probabilities of searching for each given key, and ${q_0, q_1, ..., q_n}$ represent the probabilities of searching in the gaps between and outside of these keys. We have that $\sum\limits_{i=0}^n q_i + \sum\limits_{i=1}^n p_i = 1$. We also assume without loss of generality that $q_{i-1}+p_i+q_i \neq 0$ for any $i \in \{1,...,n\}$. The keys must make up the internal nodes of the tree while the gaps make up the leaves. Our goal is to construct a binary search tree such that expected cost of search is minimized. First, we re-examine an approximate solution of G{\"u}ttler, Melhorn and Schneider which was shown to have a worst case bound of $c \cdot H + 2$ where $c \geq \frac{1}{H(\frac{1}{3},\frac{2}{3})} \approx 1.08$, and $H = \sum_{i=1}^{n} p_i\cdot\lg(\frac{1}{p_i}) + \sum_{j=0}^{n} q_i\cdot\lg(\frac{1}{q_j})$ is the entropy of the distribution. We give an improved worst case bound on the heuristic of $H+4$. Next, we examine the optimum binary search tree problem under a model of external memory. We use the Hierarchical Memory Model of Aggarwal et al. The model has an unlimited number of registers, $R_1, R_2, ...$ each with its own location in memory (a positive integer). We have a set of memory sizes $m_1, m_2, ..., m_l$ which are monotonically increasing. Each memory level has a finite size except $m_l$ which we assume has infinite size. Each memory level has an associated cost of access $c_1, c_2, ..., c_l$. We assume that $c_1 < c_2 < ... < c_l$. We propose two approximate solutions which run in $O(n)$ time where $n$ is the number of words in our data set. Using these methods, we improve upon a bound given in Thite's 2001 thesis under the related HMM$_2$ model in the approximate setting. We also examine the related problem of binary trees on multisets of probabilities where keys are unordered and we do not differentiate between which probabilities must be leaves, and which must be internal nodes. We provide a simple $O(n \lg(n))$ algorithm that is within an additive $\frac{n+1}{2n}$ of optimal on a multiset of $n$ keys.

\cleardoublepage
%\newpage

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

First and foremost, I would like to thank my supervisor Ian Munro for guiding me during my graduate studies here at the University of Waterloo. Learning from him was a great privilege. I will fondly remember our talks about data structures, algorithms, and life in general.

I would also like to thank Mordecai Golin. Conversations with Mordecai significantly helped in the simplification of the proof in \ref{MSSolvesOpt} and helped shape Chapter \ref{An Improved Bound for the Modified Minimum Entropy Heuristic}. 

I would also like to thank my lab mate Alexandre Daigle for thoroughly proof reading my first thesis draft, and spending countless hours working on whiteboard proofs with me. Also thanks to colleagues Hisham El-Zein, Dimitrios Skrepetos and Simon Pratt for working on problems with me at various times throughout my studies.    

Finally, thanks to my committee members Anna Lubiw and Eric Blais for their constructive criticism of my work.


\cleardoublepage
%\newpage

% D E D I C A T I O N
% -------------------

\begin{center}\textbf{Dedication}\end{center}

This is dedicated to my partner Melissa.
\cleardoublepage
%\newpage

% T A B L E   O F   C O N T E N T S
% ---------------------------------
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection
%\newpage

% L I S T   O F   T A B L E S
% ---------------------------
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page
%\newpage

% L I S T   O F   S Y M B O L S
% -----------------------------
% To include a Nomenclature section
% \addcontentsline{toc}{chapter}{\textbf{Nomenclature}}
% \renewcommand{\nomname}{Nomenclature}
% \printglossary
% \cleardoublepage
% \phantomsection % allows hyperref to link to the correct page
% \newpage

% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

